Comparing probabilistic predictive models applied to football

### 2015-05-22 template.tex
mexi na introducao e nos resultados

\documentclass[journal,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi} 
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes,calc,positioning}
\tdplotsetmaincoords{60}{60}
\usepackage{amsmath}
\usepackage{graphicx,subcaption}

\usepackage{amssymb}

\usepackage{color}
\newcommand{\red}[1]{\textbf{\color{red} ***ATT*** #1}}

\usepackage{changes}
\definechangesauthor[name={Rafael}, color=blue]{raf}
\setremarkmarkup{(#2)}

@@

The predictions of the complex models used here as benchmarks are published on Internet websites for every matchday, and are treated by us as black-box models since we are not specifically interested in how they were built or the assumptions on which they are based.
We know that both models are based on the bivariate Poisson or Holgate distributions, and are able to predict the final score of a match given explanatory variables.\footnote{One of the authors is responsible for one of these models and the other was described in a dissertation, see \cite{arruda2000}.}
Both models output three probabilities for each match: the probability that the home team wins, the probability that the away team wins, and the probability that there is a draw.
Because both models are based on an elaborate parametric assumptions, they are suitable for the comparison we are proposing: simple versus complex models.\footnote{To be precise we would have to define what is a ``simple'' model and what is a ``complex'' model. Based on the opening quotations, one may say that the number of parameters defines the complexity of a model. This is not the exactly the definition we will use since we consider other criteria, like the interpretation of explanatory variables and mathematical structure of the statistical model.}

@@

The black-box models used as benchmarks make their predictions available at \url{chancedegol.uol.com.br} and \url{www.previsaoesportiva.com.br}.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{ccccc}

@@

In order to have a deeper understatement about the probabilities given by each method,
Figure \ref{fig::probMandante} displays the estimated conditional probability that the home team wins given there is no draw.
All models assign higher probabilities to the home team, a fact that illustrates that
they capture \red{?the well known fact?} that it is easier for a team to win at home. 
Morever, the outputs from the methods introduced in this paper
are less dispersed than those from the other models used in the literature.

\begin{figure}[H]
  \centering
\includegraphics[page=11,scale=0.4]{futebolComparacaoModelosForPaper.pdf}
  \caption{Conditional probability that the home team wins given there is no draw. Horizontal line indicates a 50\%
  probability.}
  \label{fig::probMandante}
\end{figure}

@@

This work thus indicates that while there has been a boom in the literature with the proposal of several fancy and involved models, especially in the era of big data, simple models can many times do as well as their complex counterparts. 
This is often unnoticed by users of such methods, which leads to the overuse of models that are hard to interpret and to implement, and that also require one to keep track of several covariates.
We would call this phenomenon the {\it king's nudity}\footnote{In allusion to Anderson's {\it The Emperor's New Clothes}, \cite{emperor}: ``But he isn't wearing anything at all!''.}, that is not, of course, always observed.
In this spirit, since we have explored only the prediction of football matches, we would like to study if the the same extends to other fields such as, for instance, \red{exemplos}

@@

%The most useful statement of the principle for scientists is
%"when you have two competing theories that make exactly the same predictions, the simpler one is the better."

%What to infer from known recorded measurements about the values of yet unrecorded measurements is the responsibility of each knowing person.
%There is general agreement in evaluating the evidence that the measured observations provide for yet unmeasured experiences.
%There have been controversies among very respected scientists regarding what conclusions are to be drawn from observations.

%\acknowledgments{Acknowledgments}

%Main text.

%\authorcontributions{Author Contributions}

%Main text.

%\conflictofinterests{Conflicts of Interest}

%State any potential conflicts of interest here or ``The authors declare no conflict of interest''.

@@

@@

>>>>

\documentclass[journal,article,accept,moreauthors,pdftex,12pt,a4paper]{mdpi}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes,calc,positioning}
\tdplotsetmaincoords{60}{60}
\usepackage{amsmath}
\usepackage{graphicx,subcaption}
\usepackage{amssymb}
\usepackage{color}
\usepackage{float}

\newcommand{\red}[1]{\textbf{\color{red} ***ATT*** #1}}
\newcommand{\green}[1]{\textbf{\color{green} ***ATT*** #1}}

\usepackage{changes}
\definechangesauthor[name={Rafael}, color=blue]{raf}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Luis}, color= magenta]{luis}
\setremarkmarkup{(#2)}

@@

\added[id = {luis}]{In the statistical literature, football
prediction models can be divided in two main approaches: (i)
modeling the number of goals \citep{Maher82, Dixon97, Karlis2003}
and (ii) modeling the final outcome (win, draw or lose) of the
matches \citep{Forrest2000, Koning2000, Brillinger2008,
Brillinger2009}. For a discussion of these two approaches see
\cite{Goddard2005}. In this work, we are going to consider as
benchmarcks two model pertaining to the former approach, which has
been the most applied one.}

The predictions of the \added[id = {luis}]{two} \deleted[id = {luis}]{complex models used here as} benchmarks \added[id = {luis}]{models} are published on Internet websites for every matchday, and are treated by us as black-box models \added[id = {luis}]{.} \deleted[id = {luis}, remark = {como a seguir a gente comenta um pouco sobre as suposicoes dos modelos, talvez seja melhor omitir isto.}]{since we are not interested in how they were built or the assumptions on which they are based}. We know that \footnote{One of the authors is responsible for one of these models and the other was described in a dissertation.} both models are based \added[id = {luis}]{on Holgate bivariate Poisson regression considering as explanatory variables the home field advantage, and attack and defense strength for each team.} \added[id = {luis}]{Using} both models, \added[id = luis]{the probabilities of all possible outcomes (home team win, draw, away team win) can be calculated for each match.} \deleted[id = luis]{output three probabilities for each match: the probability that the home team wins, the probability that the away team wins, and the probability that there is a draw.}
Because both models are based on an elaborate parametric assumptions, they are suitable for the comparison we are proposing: simple versus complex models.\footnote{To be precise we would have to define what is a ``simple'' model and what is a ``complex'' model. Based on the opening quotations, one may say that the number of parameters defines the complexity of a model. This is not the exactly the definition we will use since we consider other criteria, like the interpretation of explanatory variables and mathematical structure of the statistical model.}

\added[id = {luis}]{Although the benchmark models are considered
here complex ones, there exists in the literature even more complex
models, hindering the estimation and prediction process. Among those
very complex models we can cite those based on modeling the match as
a stochastic process evolving through time \citep{Dixon98, Volf2009,
Titman2015}, those allowing for the team performance parameters to
change along the season \citep{Rue2000, Crowder2002, Owen2011,
Koopman2015} and those modeling dependence between number of goals
by means of bivariate count distribution \citep{Dixon97, Karlis2003,
McHale2007, McHale2011}.}

@@

\deleted[id = {luis}]{The black-box models used as benchmarks make their predictions available at \url{chancedegol.uol.com.br} and \url{www.previsaoesportiva.com.br}.}
\added[id = {luis}]{In the next subsections, we present the probabilistic justification for the three proposed models and present the predictive
distribution for the outcomes of an upcoming match of A (home team) versus B (away team). The main idea is to combine the past performance of both teams in different ways in order to predict the outcome of the match. The predictions of the two black-box models used as benchmarks were published just before each match-day at \url{http://chancedegol.uol.com.br} and \url{http://www.previsaoesportiva.com.br}.}

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{ccccc}

@@

In order to have a deeper understatement about the probabilities
given by each method, Figure \ref{fig::probMandante} displays the
estimated conditional probability that the home team wins given
there is no draw. \added[id = luis]{All models assign higher
probabilities to the home team win, a fact that illustrates that
they capture well the known fact of home field advantage in soccer
as well as in other sports \citep{Pollard86, Clarke95, Nevill99}.}
Morever, the outputs from the methods introduced in this paper are
less dispersed than those from the other models used in the
literature.

%\begin{figure}[H] \centering
%\includegraphics[page=11,scale=0.4]{futebolComparacaoModelosForPaper.pdf}
%\caption{Conditional probability that the home team wins given there
%is no draw. Horizontal line indicates a 50\% probability.}
%\label{fig::probMandante}
%\end{figure}

@@

This work thus indicates that while there has been a boom in the
literature with the proposal of several fancy and involved models,
especially in the era of big data, simple models can many times do
as well as their complex counterparts. This is often unnoticed by
users of such methods, which leads to the overuse of models that are
hard to interpret and to implement, and that also require one to
keep track of several covariates. We would call this phenomenon the
{\it king's nudity}\footnote{In allusion to Anderson's {\it The
Emperor's New Clothes}, \cite{emperor}: ``But he isn't wearing
anything at all!''.}, that is not, of course, always observed. In
this spirit, since we have explored only the prediction of football
matches, we would like to study if the same extends to other fields
such as, for instance, \red{exemplos}

@@

The most useful statement of the principle for scientists is "when
you have two competing theories that make exactly the same
predictions, the simpler one is the better."

What to infer from known recorded measurements about the values of yet unrecorded measurements is the responsibility of each knowing person.
There is general agreement in evaluating the evidence that the measured observations provide for yet unmeasured experiences.
There have been controversies among very respected scientists regarding what conclusions are to be drawn from observations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments{Acknowledgments}

Main text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\authorcontributions{Author Contributions}

Main text.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\conflictofinterests{Conflicts of Interest}

State any potential conflicts of interest here or ``The authors declare no conflict of interest''.

@@

\bibitem{Brillinger2008}
Brillinger, D. R. Modelling game outcomes of the Brazilian 2006
series A championship as ordinal-valued. {\em Brazilian Journal of Probability and Statistics} {\bf 2008}, {\em 22}, 89-104.

\bibitem{Brillinger2009}
Brillinger, D. R. An analysis of Chinese Super League partial
results. {\em Science in China Series A} {\bf 2009}, {\em 52},
1139-1151.

@@

\bibitem{Clarke95}
Clarke, S. R.; Norman, J. N. Home ground advantage of individual clubs in English soccer. {\em Journal of the Royal Statistical Society - Series D} {\bf 1995} ,
{\em 44}, 509-521.

\bibitem{Crowder2002} Crowder, M.; Dixon, M.; Ledford, A.; Robinson, M. Dynamic modelling and prediction of English
Football League matches for betting {\em Journal of the Royal
Statistical Society - Series D} {\bf 2002}, {\em 51}, 157-168.

\bibitem{Dixon97} Dixon, M. J.; Coles, S. C. Modelling association football scores and inefficiencies in the football betting market. {\em Applied Statistics} {\bf 1997}, {\em 46}, 265-280.

\bibitem{Dixon98} Dixon, M. J.; Robinson, M. E. A Birth Process Model for Association Football Matches. {\em Journal of the Royal Statistical Society - Series D}
{\bf 1998}, {\em 47}, 523-538.

\bibitem{Forrest2000} Forrest, D.; Simmons, R. Making up the results: The work
of the football pools panel, 1963-1997. {\em Journal of the Royal Statistical Society - Series D} {\bf 2000}, {\em 49}, 253-260.

\bibitem{Goddard2005}
Goddard, J. Regression models for forecasting goals and match
results in association football. {\em International Journal of
Forecasting} {\bf 2005}, {\em 21}, 331-340.

\bibitem{Karlis2003}
Karlis, D.; Ntzoufras, I. Analysis of sports data by using bivariate
Poisson models. {\em Journal of the Royal Statistical Society - Series D} {\bf 2003}, {\em 52}, 381-393.

\bibitem{Koning2000}
Koning, R. H. . Balance in competition in Dutch soccer. {\em Journal of the Royal Statistical Society - Series D} {\bf 2000}, {\em 49}, 419-431.

\bibitem{Koopman2015}
Koopman, S. J.; Lit, Rutger. A dynamic bivariate Poisson model for
analysing and forecasting match results in the English Premier
League. {\em Journal of the Royal Statistical Society - Series A}
{\bf 2015}, {\em 178}, 167-186.

\bibitem{Maher82}
Maher, M. J. Modelling association football scores. {\em Statistica Neerlandica} {\bf 1982}, {\em 36}, 109-118.

\bibitem{McHale2007}
McHale, I.; Scarf, P. Modelling soccer matches using bivariate
discrete distributions with general dependence structure. {\em
Statistica Neerlandica} {\bf 2007}, {\em 61}, 432-445.

\bibitem{McHale2011}
McHale, I.; Scarf, P. Modelling the dependence of goals scored by
opposing teams in international soccer matches. {\em Statistical
Modelling} {\bf 2011}, {\em 11}, 219–236.

\bibitem{Nevill99}
Nevill, A. M.; Holder, R. L. Home advantage in sport. {\em Sports
Medicine} {\bf 1999}, {\em 28}, 221-236.

\bibitem{Owen2011}
Owen, A. Dynamic Bayesian forecasting models of football match
outcomes with estimation of the evolution variance parameter. {\em
IMA Journal of Management Mathematics} {\bf 2011}, {\em 22}, 99-113.

\bibitem{Pollard86}
Pollard, R. Home advantage in soccer: a retrospective analysis. {\em Journal of Sports Sciences} {\bf 1986}, {\em 4}, 237-248.

\bibitem{Rue2000}
Rue, H.; Salvesen, O. Prediction and Retrospective Analysis of
Soccer Matches in a League {\em Journal of the Royal Statistical
Society - Series D} {\bf 2000}, {\em 49}, 399-418.

\bibitem{Titman2015}
Titman, A. C.; Costain, D. A.; Ridall, P. G.; Gregory, K. Joint
modelling of goals and bookings in association football. {\em
Journal of the Royal Statistical Society - Series A} {\bf 2015},
{\em 178}, 659-683.

\bibitem{Volf2009}
Volf, P. A random point process model for the score in sport
matches. {\em IMA Journal of Management Mathematics} {\bf 2009},
{\em 20}, 121-131.
